---
title: "Exercises"
author: "Sergio Rodriguez Llana"
date: "5/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**GitHub respository:** https://github.com/SergioRLL/SRL_Omics_Techniques.git

# Exercise 1

- **Title**: Early postnatal expression data from mouse skeletal muscle stem cells
- **GEO identifier (GSE)**: GSE65927
- **Summary of the study**: Satellite cells are the primary source of stem cells for skeletal muscle growth and regeneration. Since adult stem cell maintenance involves a fine balance between intrinsic and extrinsic mechanisms, the researchers performed genome-wide chronological expression profiling to identify the transcriptomic changes involved during early postnatal growth till acquisition of satellite cell quiescence.
- **Experimental design**: Muscle samples were isolated from the trunk of Pax3GFP/+ mice. After digestion, GFP cells were purified via FACS and process for RNA extraction and hybridization on Affymetrix microarrays (Affymetrix Mouse Genome 430 2.0 Arrays).
The different ages selected for sample isolation were P1-P12-P28 (P, postnatal days), covering proliferant and quiescent satellite cells. The three stages were done in triplicate for P28 and 4 times for P1-P12, so 11 samples included in the microarray.

```{r, results='hide'}
require(GEOquery)
gse <- getGEO("GSE65927")
```

```{r}
names(gse)
gse[[1]]
esetFromGEO <- gse[[1]]
```

## Expression Matrix:

```{r}
require(Biobase)
head(exprs(esetFromGEO))
dim(exprs(esetFromGEO))
```

Looking at the Expression Matrix dimensions we see that there are 45101 rows and 11 columns. Columns correspond to different samples.

```{r}
summary(exprs(esetFromGEO))
```

It seems that the data is normalzied as the mean and the meadian has a close vaule.


## PhenoData:


```{r}
head(pData(esetFromGEO))
ph_data <- pData(esetFromGEO)
colnames(pData(esetFromGEO))
```

Now I'm going to see the number of groups of the experiment (factors):

```{r}
ph_data$title
```

We have the following factors:

- A factor that is the postnatal day: (Postnatal P1, Postnatal P12 and Postnatal P28).
- A factor that corresponds to the biologial replicate: (rep1, rep2 rep3 and rep4).

Therefore we sould have a total of 12 groups. However the samples I think that do not cover all the posibilites.

## Plots:

```{r}
boxplot(exprs(esetFromGEO))
hist(exprs(esetFromGEO), main = "Histogram of expression values", xlab = "Expression values")

```

We can observe that the data have similar variance. The most expressed values are arround 5.


